### 2025-01-27 论文/PAPER.tex
论文

%% 美赛模板：正文部分

\documentclass[12pt]{article}  % 官方要求字号不小于 12 号，此处选择 12 号字体

% 本模板不需要填写年份，以当前电脑时间自动生成
% 请在以下的方括号中填写队伍控制号
\usepackage[2506135]{easymcm}  % 载入 EasyMCM 模板文件
\problem{C}  % 请在此处填写题号
\usepackage{mathptmx}  % 这是 Times 字体，中规中矩 
%\usepackage{mathpazo}  % 这是 COMAP 官方杂志采用的更好看的 Palatino 字体，可替代以上的 mathptmx 宏包
\usepackage{amsmath}
\usepackage{tabularray}
%拉取的一些宏包
\usepackage{graphicx} 
%\usepackage{pdfpages}
\usepackage{longtable}
%\usepackage{tabu}
\usepackage{threeparttable} %支持在表格中使用 \tnote 命令添加脚注，并自动编号
\usepackage{listings} %在文档中插入代码
%\usepackage{paralist}%创建紧凑的列表，节省垂直空间
%\usepackage{setspace} %调整文档的行间距
%\usepackage{booktabs}%用于创建更美观的表格，提供了高质量的表格线条命令，如 \toprule、\midrule 和 \bottomrule
%\usepackage{makecell} %创建多行单元格，支持复杂的表格布局
\usepackage{amssymb} %提供了额外的数学符号，通常与 amsmath 宏包一起使用
\usepackage{appendix} %管理附录部分，自动处理附录的编号和格式
\newcommand{\upcite}[1]{\textsuperscript{\textsuperscript{\cite{#1}}}}

\title{Together, Individuals Make a Difference}  % 标题

% 如需要修改题头（默认为 MCM/ICM），请使用以下命令（此处修改为 MCM）
%\renewcommand{\contest}{MCM}

% 文档开始
\newtheorem{keywords}{Keywords}
\begin{document}

% 此处填写摘要内容
\begin{abstract}
    Here is the abstract of your paper.

    Firstly, that is ...

    Secondly, that is ...

    Finally, that is ...

$F(\omega) = \int_{-\infty}^{\infty} f(t) e^{-i \omega t} \, dt$

$f(t) = \frac{1}{2\pi} \int_{-\infty}^{\infty} F(\omega) e^{i \omega t} \, d\omega$

    % 美赛论文中无需注明关键字。若您一定要使用，
    % 请将以下两行的注释号 '%' 去除，以使其生效
    % \vspace{5pt}
    % \textbf{Keywords}: MATLAB, mathematics, LaTeX.

    \textbf{PCA}\\

    % 关键字Keywords
    \vspace{5pt}
    \textbf{Keywords}: \textbf{ A, B, C,  }

\end{abstract}

\maketitle  % 生成 Summary Sheet
\tableofcontents  % 生成目录

% 正文开始
\section{Introduction}
\subsection{Problem Background}
During the 2024 Paris Summer Olympics, fans not only paid attention to the individual events but also showed significant interest in the the ranking of countries on the overall medal count and  gold medal table. Ultimately, the United States topped the medal table with a total of 126 medals, while China and the United States shared the top position on the gold medal table, each securing 40 golds. The host country, France, ranked fifth on the gold medal table with 16 gold medals, but finished fourth in the total medal count. Great Britain ranked seventh on the gold medal table with 14 golds, while securing third place in the overall medal count. Despite the prominence of the leading countries, the medal achievements of other nations also attracted attention. For example, Albania, Cape Verde, Dominica, and Saint Lucia each won their first-ever Olympic medal in this edition of the Games, with Dominica and Saint Lucia each earning another gold medal. However, over 60 countries have yet to win a medal in the history of the Olympics.

\subsection{Restatement of Problem}
While predictions of the final medal counts at the Olympics are common, such forecasts are generally not based on historical medal tables. Instead, they are typically made before the start of the upcoming Olympic Games, once the list of competing athletes is known.To clarify the task, the problem is restated as follows:

\begin{itemize}
	\setlength{\parsep}{0ex} %段落间距
	\setlength{\topsep}{2ex} %列表到上下文的垂直距离
	\setlength{\itemsep}{1ex} %条目间距
	\item Develop a model for each country’s medal count, which should at least include the number of gold medals and the total medal count, and estimate the uncertainty/precision of the model,also evaluate its performance.
		\begin{itemize}
		\item[1)]
	Using the model, predict the medal standings for the 2028 Los Angeles Summer Olympics, including prediction intervals for all outcomes. Based on the model’s predictions, identify which countries are most likely to improve their performance and which countries are expected to perform worse than in the 2024.
	\end{itemize}
	\begin{itemize}
		\item[2)]
		The developed model should include countries that have not yet won any medals, while also predicting the number of countries likely to win their first medal in the upcoming Olympics and provide odds for this estimate.
	\end{itemize}
		\begin{itemize}
		\item[3)]
	The model should also consider the number and types of events in each specific Olympic Games, exploring the relationship between the events and the number of medals won by each country. For each country, determine which events are most crucial and why, as well as the impact of the country’s selected events on the results.
	\end{itemize}

	\item In contrast to athletes, who must change their citizenship in order to represent different countries in competitions, coaches do not need to change their citizenship to coach, which makes it easier for them to move to other countries. As a result, the “Great Coach” effect may arise. For example, Lang Ping has coached the volleyball teams of both China and the U.S., leading both to championships, while Bela Karolyi coached the Romanian and U.S. women’s gymnastics teams, achieving great success. The task is to identify evidence in the data that may suggest changes driven by the "Great Coach" effect and estimate its impact on the medal count. Select three countries, determine the events in which they should consider investing in "outstanding" coaches, and estimate the potential impact of such investments.
	\item Explain the unique insights regarding Olympic medal counts that are included in the developed model and how these insights can provide valuable information to the National Olympic Committees of various countries.

\end{itemize}

\subsection{Our work}
We do such things ...

%图片示例（water.png）
%\begin{figure}[htbp]
%	\centering
%	\includegraphics[width=.8\textwidth]{water.png}
%	\caption{The result of Model 2}\label{fig:result}
%\end{figure}
\begin{enumerate}[\bfseries 1.]
    \item We do ...
    \item We do ...
    \item We do ...
\end{enumerate}
\section{Question Preparation}
\subsection{Assumptions}

\subsection{Notations}

The primary notations used in this paper are listed in Table \ref{tb:notation}.

% 三线表示例
\begin{table}[!htbp]
\begin{center}
\caption{Notations}
\begin{tabular}{cl}
	\toprule
	\multicolumn{1}{m{3cm}}{\centering Symbol}
	&\multicolumn{1}{m{8cm}}{\centering Definition}\\
	\midrule
	$C_{i}$&the first one\\
	$S_{i}$&the second one\\
	$G$&the second one\\
	$V$&the second one\\
	$B$&the second one\\
	$M$&the second one\\
	$A$&the second one\\
	$D$&the second one\\
	$E$&the second one\\
%	$\alpha$ &the last one\\
    $\gamma$&Sport Advantagen Coefficient\\
  	$ m_{c_{i}}$ &the second one\\
	$ m_{c_{i} s_{j}}$ &the second one\\
    $g_{i}$&the second one\\
    $\xi$&The Probability of First Medal\\
	\bottomrule
\end{tabular}\label{tb:notation}
\end{center}
\end{table}

\section{Data Preprocessing}
\subsection{Basic Data Preprocessing}

\begin{equation}
G_{i} \cap G_{j}=\emptyset, \bigcup_{i=1}^{5} G_{i}=X
\end{equation}
\begin{figure}[htbp]
	\centering
	\includegraphics[width=12cm]{img/Level1.png}
	\caption{Scatter plot of national level classification (based on Kmeans++clustering algorithm)}
	\label{fig:aa}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=12cm]{img/Level2.png}
	\caption{aa}
	\label{fig:aa}
\end{figure}

\subsection{Data Mining}
\subsubsection{Athlete Service Status}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=12cm]{img/Number.png}
	\caption{Number of New and Leave Athletes in Each Year}
	\label{fig:aa}
\end{figure}

% 子图（多图并列）示例，更多用法请参考 subfigure 宏包文档
% 如果您只希望几张图并列，不需要额外的 caption，那么在 figure 环境中
% 连续插入总宽度不超过 \textwidth 的多个 \includegraphics 命令即可
\begin{figure}[htbp]
	\centering
	\begin{subfigure}[b]{.55\textwidth}
		\includegraphics[width=\textwidth]{img/Times.png}
		\caption{Distribution of Serve Times}\label{subfig:left}
	\end{subfigure}
	\begin{subfigure}[b]{.4\textwidth}
		\includegraphics[width=\textwidth]{img/Years.png}
		\caption{Distribution of Serve Years}\label{subfig:right}
	\end{subfigure}
	\caption{Two images}\label{fig:subfigures}
\end{figure}

\subsubsection{Distribution of Countries' Strength Sports}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=12cm]{img/Percentage.jpg}
	\caption{Percentage of Medals of Different Sports in China}
	\label{fig:aa}
\end{figure}

\section{Task1:Medal Prediction Model Based on LightGBM}
\subsection{Medal Standings}
The detail can be described by equation \eqref{eq1}:\autoref{eq1}:

\begin{equation}
\gamma = \frac{m_{c_i s_j}}{\sum m_{c_i s_j}}
\end{equation}

%Table 2 :变量名表 Variable Name
\begin{longtblr}[
	caption = {Variable Name},
	]{
		vline{1-4} = {1-10}{},
		hline{1-11} = {1-3}{},
	}
	Variable Name                                                    & \textit{Code}                                                & Definition                                                                      &  \\
	Whether~Host Country                                             & \textit{is host}                                             & {Whether the country is the host(1 for host,0 for\\~non-host)}                  &  \\
	{Medal Expectation\\Increment *Personnel\\Expectation lncrement} & {\textit{medal\_increment *}\\\textit{personnel\_increment}} & {Product of medal\\expectation increment and\\personnel expectation\\increment} &  \\
	{Sport Advantage\\Coefficient}                                 & \textit{sport\_adv}                                           & {Advantage coefficient of\\a specific sport}                                  &  \\
	Country Level                                                    & \textit{country\_lvl}                                        & {The level of the country in the competition\\(ordered by rank)}                &  \\
	{Project Medal\\Expectation /Project\\Personnel Expection~}                         & \textit{sport\_medal \_per\_ person}                          & {Ratio of sport medals to projected personnel\\for a specific sport}          &  \\
	Gold Medal Probability                                           & \textit{gold\_prob}                                          & {Probability of an athlete~\\winning a gold medal}                              &  \\
	{Silver Medal\\Probability}                                      & \textit{silver\_prob}                                        & Probability of an athlete~ winning a silver medal                               &  \\
	{Bronze Medal\\Probability}                                      & \textit{silver\_prob}                                        & Probability of an athletewinning a bronze medal                                 &  \\
	No Medal Probability                                             & \textit{no\_medal\_probe}                                    & Probability of an athlete winning no medal                                      &  \\
	&                                                              &                                                                                 &  \\
	&                                                              &                                                                                 &  
\end{longtblr}

%奖牌榜预测图

\begin{figure}[htbp]
	\centering
	\includegraphics[width=12cm]{img/Predict.png}
	\caption{Medal count prediction}
	\label{fig:aa}
\end{figure}

% \usepackage{longtable}

\begin{longtable}{l|c|c|c|c|c|c|c|c|c|c}
\caption{Countries' Medal Count Prediction(part)}\\
	Country              & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}G\_\\pes.\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}S\_\\pes.\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}B\_\\pes.\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}G\_\\opt.\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}S\_\\opt.\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}B\_\\opt.\end{tabular}} & \multicolumn{1}{l|}{Gold} & \multicolumn{1}{l|}{Silver} & \multicolumn{1}{l|}{Bronze} & \multicolumn{1}{l}{Total}  \endfirsthead 
	\hline
	United States        & 47                                                                     & 36                                                                     & 23                                                                     & 53                                                                     & 41                                                                     & 25                                                                     & 51                        & 40                          & 25                          & 117                        \\ 
	\hline
	China                & 35                                                                     & 35                                                                     & 15                                                                     & 46                                                                     & 38                                                                     & 20                                                                     & 40                        & 36                          & 17                          & 95                         \\ 
	\hline
	Australia            & 28                                                                     & 20                                                                     & 22                                                                     & 33                                                                     & 24                                                                     & 24                                                                     & 29                        & 22                          & 23                          & 71                         \\ 
	\hline
	France               & 23                                                                     & 19                                                                     & 19                                                                     & 27                                                                     & 20                                                                     & 22                                                                     & 24                        & 19                          & 21                          & 64                         \\ 
	\hline
	Germany              & 18                                                                     & 24                                                                     & 14                                                                     & 23                                                                     & 26                                                                     & 16                                                                     & 21                        & 25                          & 15                          & 61                         \\ 
	\hline
	United Kingdom       & 20                                                                     & 19                                                                     & 14                                                                     & 26                                                                     & 24                                                                     & 15                                                                     & 25                        & 21                          & 15                          & 61                         \\ 
	\hline
	Japan                & 16                                                                     & 25                                                                     & 14                                                                     & 19                                                                     & 27                                                                     & 16                                                                     & 17                        & 26                          & 16                          & 58                         \\ 
	\hline
	Italy                & 15                                                                     & 20                                                                     & 15                                                                     & 17                                                                     & 20                                                                     & 22                                                                     & 16                        & 20                          & 16                          & 53                         \\ 
	\hline
	Spain                & 21                                                                     & 10                                                                     & 13                                                                     & 30                                                                     & 12                                                                     & 14                                                                     & 25                        & 12                          & 13                          & 50                         \\ 
	\hline
	Netherlands          & 24                                                                     & 6                                                                      & 13                                                                     & 29                                                                     & 9                                                                      & 15                                                                     & 26                        & 8                           & 15                          & 47                         \\ 
	\hline
	New Zealand          & 20                                                                     & 13                                                                     & 9                                                                      & 22                                                                     & 15                                                                     & 10                                                                     & 22                        & 14                          & 9                           & 45                         \\ 
	\hline
	...                  & ..                                                                     & ..                                                                     & ..                                                                     & ..                                                                     & ..                                                                     & ..                                                                     & ..                        & ..                          & ..                          & ..                         \\
	\multicolumn{1}{l}{} & \multicolumn{1}{c}{}                                                   & \multicolumn{1}{c}{}                                                   & \multicolumn{1}{c}{}                                                   & \multicolumn{1}{c}{}                                                   & \multicolumn{1}{c}{}                                                   & \multicolumn{1}{c}{}                                                   & \multicolumn{1}{c}{}      & \multicolumn{1}{c}{}        & \multicolumn{1}{c}{}        &                           
\end{longtable}

\subsection{Countries that Win Their First Medal}

%首奖可能性图

\begin{figure}[htbp]
	\centering
	\includegraphics[width=16cm]{img/First.png}
	\caption{The Probability of First Medal in 2028}
	\label{fig:aa}
\end{figure}

\subsection{Events and Medal Counts by Countries}

%三个国家擅长项目的图
\begin{figure}[htbp]
	\centering
	\begin{subfigure}[b]{.3\textwidth}
		\includegraphics[width=\textwidth]{img/Table Tennis.png}
		\caption{Heatmap of Correlation for Table Tennis in China}\label{subfig:1}
	\end{subfigure}
	\hfill 
	\begin{subfigure}[b]{.33\textwidth}
		\includegraphics[width=\textwidth]{img/Jamaica.png}
		\caption{ Heatmap of Correlation for Athletics in Jamaica}\label{subfig:2}
	\end{subfigure}
	\hfill 
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/New Zealand.png}
		\caption{Heatmap of Correlation for Rowing in New Zealand}\label{subfig:3}
	\end{subfigure}
	\caption{Three images}\label{fig:subfigures}
\end{figure}

\section{Task2:}

The results are shown in Figure \ref{fig:result}, where $t$ denotes the time in seconds, and $c$ refers to the concentration of water in the boiler.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=16cm]{img/Volleyball.png}
	\caption{Medal Count of Women's Volleyball Teams(CHN vs USA) Over Time}
	\label{fig:aa}
\end{figure}

%三个国家某项目衰弱 需要教练图
\begin{figure}[htbp]
	\centering
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/Decline1.png}
		\caption{France}\label{subfig:1}
	\end{subfigure}
	\hfill 
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/Decline2.png}
		\caption{Sweden}\label{subfig:2}
	\end{subfigure}
	\hfill 
	\begin{subfigure}[b]{.32\textwidth}
		\includegraphics[width=\textwidth]{img/Decline3.png}
		\caption{Germany}\label{subfig:3}
	\end{subfigure}
	\caption{Three images}\label{fig:subfigures}
\end{figure}

\clearpage

%引用图像示例Figure \ref{fig:subfigures}
%Figure \ref{fig:subfigures} gives an example of subfigures. Figure \ref{subfig:left} is on the left, and Figure \ref{subfig:right} is on the right.

% \usepackage{longtable}

\begin{longtable}{l|l|l|l|l} 
	\caption{Model Performances(LightGBM)}\\ 
	\hline
	\multicolumn{1}{c|}{\textbf{Model}} & \multicolumn{1}{c|}{\textbf{MSE}} & \multicolumn{1}{c|}{\textbf{RMSE}} & \multicolumn{1}{c|}{\textbf{MAE}} & \multicolumn{1}{c}{\(\mathbf{R}^2\)}  \endfirsthead 
	\hline
	Gold Prediction                     & 0.0235                            & 0.180                              & 0.030                             & 0.690                              \\ 
	\hline
	Silver Prediction                   & 0.0239                            & 0.185                              & 0.033                             & 0.625                              \\ 
	\hline
	Bronze Prediction                   & 0.0231                            & 0.187                              & 0.034                             & 0.606                              \\
	\hline
\end{longtable}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=16cm]{img/Performance.png}
	\caption{Model Performance Radar Chart}
	\label{fig:aa}
\end{figure}

\section{Task3}

%公式代码集合

\begin{equation}
	\mathcal{L} = \sum_{i=1}^{n} L(y_i, \hat{y}_i^{(m-1)} + \eta f_m(x_i)) + \Omega(f_m)
\end{equation}

$y_i$\\
$\hat{y}_i^{(m-1)}$\\
$\eta f_m(x_i)$\\
$\Omega(f_m)$\\

\begin{equation}
	\kappa = 0.8 \times \frac{\xi - \overline{\xi}}{\sigma}
\end{equation}

$\overline{\xi}$
$\sigma$

\begin{equation}
	d_i = R(X_i) - R(Y_i)
\end{equation}

\begin{equation}
	R_s = 1 - \frac{6 \sum d_i^2}{n(n^2 - 1)}
\end{equation}

$d_i$
$R_s$

\begin{equation}
	P(c_k \mid X) = \frac{P(X \mid c_k) P(c_k)}{P(X)}
\end{equation}

\begin{equation}
	C_t^+ = \max(0, C_{t-1}^+ + (X_t - \mu - k))
\end{equation}
$C_t^+ $

	\begin{equation}
		\phi_i(f) = \sum_{S \subseteq N \setminus \{i\}} \frac{|S|!(|N| - |S| - 1)!}{|N|!} [f(S \cup \{i\}) - f(S)]
	\end{equation}

$\phi_i(f)$

\section{Sensitivity Analysis}

\section{Model Evaluation}
\subsection{Strengths}
1.

2.

3.

4. 
\subsection{Weaknesses}
1.

\section{Conclusion}
\begin{itemize}
	\item Task1

	 \qquad AA\\
	bbb

	\item Task2\\
	\qquad AAA
	\item Task3\\
	\qquad AAA

	\item Task4\\
	\qquad AAA

\end{itemize}

% 参考文献，此处以 MLA 引用格式为例
\begin{thebibliography}{99}
\bibitem{1} Einstein, A., Podolsky, B., \& Rosen, N. (1935). Can quantum-mechanical description of physical reality be considered complete?. \emph{Physical review}, 47(10), 777.
\bibitem{2} \emph{A simple, easy \LaTeX\ template for MCM/ICM: EasyMCM}. (2018). Retrieved December 1, 2019, from\url{https://www.cnblogs.com/xjtu-blacksmith/p/easymcm.html}
\end{thebibliography}

% 以下为附录内容
% 如您的论文中不需要附录，请自行删除
\begin{subappendices}  % 附录环境

\section{Appendix A: Countries' Medal Count Prediction}

\begin{longtable}{|l|c|c|c|c|c|c|c|c|c|c|}
	\caption{Countries' Medal Count Prediction}\\
	\hline
	Country                                                        & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}G\_\\pes.\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}S\_\\pes.\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}B\_\\pes.\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}G\_\\opt.\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}S\_\\opt.\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}B\_\\opt.\end{tabular}} & \multicolumn{1}{l|}{Gold} & \multicolumn{1}{l|}{Silver} & \multicolumn{1}{l|}{Bronze} & \multicolumn{1}{l|}{Total}  \endfirsthead 
	\hline
	United States                                                  & 47                                                                     & 36                                                                     & 23                                                                     & 53                                                                     & 41                                                                     & 25                                                                     & 51                        & 40                          & 25                          & 117                         \\ 
	\hline
	China                                                          & 35                                                                     & 35                                                                     & 15                                                                     & 46                                                                     & 38                                                                     & 20                                                                     & 40                        & 36                          & 17                          & 95                          \\ 
	\hline
	Australia                                                      & 28                                                                     & 20                                                                     & 22                                                                     & 33                                                                     & 24                                                                     & 24                                                                     & 29                        & 22                          & 23                          & 71                          \\ 
	\hline
	France                                                         & 23                                                                     & 19                                                                     & 19                                                                     & 27                                                                     & 20                                                                     & 22                                                                     & 24                        & 19                          & 21                          & 64                          \\ 
	\hline
	Germany                                                        & 18                                                                     & 24                                                                     & 14                                                                     & 23                                                                     & 26                                                                     & 16                                                                     & 21                        & 25                          & 15                          & 61                          \\ 
	\hline
	United Kingdom                                                 & 20                                                                     & 19                                                                     & 14                                                                     & 26                                                                     & 24                                                                     & 15                                                                     & 25                        & 21                          & 15                          & 61                          \\ 
	\hline
	Japan                                                          & 16                                                                     & 25                                                                     & 14                                                                     & 19                                                                     & 27                                                                     & 16                                                                     & 17                        & 26                          & 16                          & 58                          \\ 
	\hline
	Italy                                                          & 15                                                                     & 20                                                                     & 15                                                                     & 17                                                                     & 20                                                                     & 22                                                                     & 16                        & 20                          & 16                          & 53                          \\ 
	\hline
	Spain                                                          & 21                                                                     & 10                                                                     & 13                                                                     & 30                                                                     & 12                                                                     & 14                                                                     & 25                        & 12                          & 13                          & 50                          \\ 
	\hline
	Netherlands                                                    & 24                                                                     & 6                                                                      & 13                                                                     & 29                                                                     & 9                                                                      & 15                                                                     & 26                        & 8                           & 15                          & 47                          \\ 
	\hline
	New Zealand                                                    & 20                                                                     & 13                                                                     & 9                                                                      & 22                                                                     & 15                                                                     & 10                                                                     & 22                        & 14                          & 9                           & 45                          \\ 
	\hline
	Canada                                                         & 15                                                                     & 10                                                                     & 16                                                                     & 19                                                                     & 10                                                                     & 18                                                                     & 17                        & 10                          & 16                          & 43                          \\ 
	\hline
	Brazil                                                         & 8                                                                      & 10                                                                     & 13                                                                     & 8                                                                      & 11                                                                     & 15                                                                     & 8                         & 10                          & 14                          & 32                          \\ 
	\hline
	Belgium                                                        & 6                                                                      & 8                                                                      & 7                                                                      & 15                                                                     & 10                                                                     & 8                                                                      & 11                        & 8                           & 7                           & 26                          \\ 
	\hline
	Hungary                                                        & 7                                                                      & 6                                                                      & 8                                                                      & 8                                                                      & 8                                                                      & 9                                                                      & 8                         & 8                           & 9                           & 25                          \\ 
	\hline
	Poland                                                         & 10                                                                     & 7                                                                      & 6                                                                      & 12                                                                     & 10                                                                     & 8                                                                      & 11                        & 9                           & 6                           & 25                          \\ 
	\hline
	Ireland                                                        & 6                                                                      & 10                                                                     & 6                                                                      & 7                                                                      & 11                                                                     & 6                                                                      & 6                         & 10                          & 6                           & 23                          \\ 
	\hline
	Argentina                                                      & 4                                                                      & 8                                                                      & 7                                                                      & 5                                                                      & 10                                                                     & 9                                                                      & 4                         & 9                           & 8                           & 22                          \\ 
	\hline
	Denmark                                                        & 10                                                                     & 7                                                                      & 5                                                                      & 12                                                                     & 8                                                                      & 8                                                                      & 10                        & 8                           & 6                           & 22                          \\ 
	\hline
	Ukraine                                                        & 3                                                                      & 7                                                                      & 8                                                                      & 5                                                                      & 8                                                                      & 9                                                                      & 4                         & 8                           & 8                           & 21                          \\ 
	\hline
	South Korea                                                    & 10                                                                     & 5                                                                      & 4                                                                      & 12                                                                     & 6                                                                      & 4                                                                      & 11                        & 6                           & 4                           & 21                          \\ 
	\hline
	Romania                                                        & 10                                                                     & 7                                                                      & 3                                                                      & 12                                                                     & 8                                                                      & 3                                                                      & 12                        & 7                           & 3                           & 20                          \\ 
	\hline
	Norway                                                         & 11                                                                     & 1                                                                      & 4                                                                      & 11                                                                     & 3                                                                      & 6                                                                      & 11                        & 2                           & 6                           & 19                          \\ 
	\hline
	South Africa                                                   & 6                                                                      & 4                                                                      & 4                                                                      & 8                                                                      & 8                                                                      & 5                                                                      & 8                         & 7                           & 5                           & 18                          \\ 
	\hline
	Slovenia                                                       & 4                                                                      & 3                                                                      & 8                                                                      & 6                                                                      & 3                                                                      & 8                                                                      & 4                         & 3                           & 8                           & 17                          \\ 
	\hline
	Kenya                                                          & 5                                                                      & 4                                                                      & 4                                                                      & 5                                                                      & 7                                                                      & 5                                                                      & 5                         & 7                           & 4                           & 16                          \\ 
	\hline
	Serbia                                                         & 4                                                                      & 5                                                                      & 6                                                                      & 5                                                                      & 9                                                                      & 7                                                                      & 4                         & 8                           & 6                           & 16                          \\ 
	\hline
	Mexico                                                         & 4                                                                      & 4                                                                      & 3                                                                      & 6                                                                      & 4                                                                      & 5                                                                      & 5                         & 4                           & 4                           & 15                          \\ 
	\hline
	India                                                          & 1                                                                      & 1                                                                      & 4                                                                      & 8                                                                      & 4                                                                      & 5                                                                      & 7                         & 3                           & 4                           & 15                          \\ 
	\hline
	Greece                                                         & 3                                                                      & 4                                                                      & 2                                                                      & 7                                                                      & 4                                                                      & 3                                                                      & 7                         & 4                           & 3                           & 14                          \\ 
	\hline
	Switzerland                                                    & 2                                                                      & 4                                                                      & 5                                                                      & 5                                                                      & 4                                                                      & 6                                                                      & 4                         & 4                           & 6                           & 14                          \\ 
	\hline
	Jamaica                                                        & 1                                                                      & 7                                                                      & 2                                                                      & 5                                                                      & 12                                                                     & 3                                                                      & 4                         & 8                           & 2                           & 13                          \\ 
	\hline
	Czech Republic                                                 & 5                                                                      & 3                                                                      & 3                                                                      & 6                                                                      & 4                                                                      & 5                                                                      & 6                         & 4                           & 4                           & 13                          \\ 
	\hline
	Uzbekistan                                                     & 4                                                                      & 2                                                                      & 2                                                                      & 5                                                                      & 3                                                                      & 2                                                                      & 4                         & 2                           & 2                           & 10                          \\ 
	\hline
	Nigeria                                                        & 4                                                                      & 2                                                                      & 4                                                                      & 4                                                                      & 3                                                                      & 4                                                                      & 4                         & 2                           & 4                           & 10                          \\ 
	\hline
	Sweden                                                         & 4                                                                      & 4                                                                      & 2                                                                      & 5                                                                      & 5                                                                      & 5                                                                      & 4                         & 4                           & 2                           & 10                          \\ 
	\hline
	Turkey                                                         & 3                                                                      & 2                                                                      & 5                                                                      & 3                                                                      & 2                                                                      & 6                                                                      & 3                         & 2                           & 5                           & 10                          \\ 
	\hline
	Finland                                                        & 2                                                                      & 6                                                                      & 2                                                                      & 2                                                                      & 6                                                                      & 2                                                                      & 2                         & 6                           & 2                           & 10                          \\ 
	\hline
	Colombia                                                       & 5                                                                      & 2                                                                      & 0                                                                      & 5                                                                      & 3                                                                      & 1                                                                      & 5                         & 3                           & 0                           & 9                           \\ 
	\hline
	Egypt                                                          & 0                                                                      & 3                                                                      & 3                                                                      & 1                                                                      & 4                                                                      & 4                                                                      & 0                         & 4                           & 4                           & 8                           \\ 
	\hline
	Kazakhstan                                                     & 1                                                                      & 2                                                                      & 2                                                                      & 4                                                                      & 4                                                                      & 2                                                                      & 3                         & 3                           & 2                           & 8                           \\ 
	\hline
	Iran                                                           & 2                                                                      & 2                                                                      & 1                                                                      & 5                                                                      & 3                                                                      & 2                                                                      & 4                         & 2                           & 1                           & 8                           \\ 
	\hline
	\begin{tabular}[c]{@{}l@{}}Dominican \\Republic\end{tabular}   & 3                                                                      & 3                                                                      & 2                                                                      & 6                                                                      & 4                                                                      & 3                                                                      & 5                         & 3                           & 3                           & 8                           \\ 
	\hline
	Portugal                                                       & 2                                                                      & 3                                                                      & 3                                                                      & 2                                                                      & 4                                                                      & 3                                                                      & 2                         & 3                           & 3                           & 8                           \\ 
	\hline
	Cuba                                                           & 2                                                                      & 2                                                                      & 1                                                                      & 3                                                                      & 2                                                                      & 2                                                                      & 3                         & 2                           & 2                           & 7                           \\ 
	\hline
	Unknown                                                        & 1                                                                      & 2                                                                      & 2                                                                      & 1                                                                      & 2                                                                      & 4                                                                      & 1                         & 2                           & 2                           & 7                           \\ 
	\hline
	Latvia                                                         & 3                                                                      & 1                                                                      & 3                                                                      & 3                                                                      & 1                                                                      & 3                                                                      & 3                         & 1                           & 3                           & 7                           \\ 
	\hline
	Ethiopia                                                       & 2                                                                      & 1                                                                      & 0                                                                      & 2                                                                      & 6                                                                      & 2                                                                      & 2                         & 2                           & 2                           & 6                           \\ 
	\hline
	Algeria                                                        & 2                                                                      & 2                                                                      & 1                                                                      & 2                                                                      & 3                                                                      & 2                                                                      & 2                         & 2                           & 1                           & 6                           \\ 
	\hline
	Morocco                                                        & 0                                                                      & 2                                                                      & 4                                                                      & 0                                                                      & 2                                                                      & 5                                                                      & 0                         & 2                           & 5                           & 6                           \\ 
	\hline
	Austria                                                        & 3                                                                      & 0                                                                      & 1                                                                      & 3                                                                      & 1                                                                      & 2                                                                      & 3                         & 0                           & 2                           & 6                           \\ 
	\hline
	Puerto Rico                                                    & 2                                                                      & 3                                                                      & 1                                                                      & 4                                                                      & 3                                                                      & 2                                                                      & 2                         & 3                           & 2                           & 6                           \\ 
	\hline
	Thailand                                                       & 1                                                                      & 2                                                                      & 1                                                                      & 2                                                                      & 2                                                                      & 2                                                                      & 2                         & 2                           & 1                           & 5                           \\ 
	\hline
	Croatia                                                        & 2                                                                      & 2                                                                      & 0                                                                      & 2                                                                      & 2                                                                      & 1                                                                      & 2                         & 2                           & 1                           & 5                           \\ 
	\hline
	Ecuador                                                        & 2                                                                      & 0                                                                      & 2                                                                      & 3                                                                      & 1                                                                      & 3                                                                      & 2                         & 0                           & 3                           & 5                           \\ 
	\hline
	Uganda                                                         & 0                                                                      & 1                                                                      & 1                                                                      & 1                                                                      & 4                                                                      & 1                                                                      & 1                         & 3                           & 1                           & 5                           \\ 
	\hline
	Angola                                                         & 0                                                                      & 2                                                                      & 0                                                                      & 1                                                                      & 3                                                                      & 1                                                                      & 0                         & 2                           & 0                           & 5                           \\ 
	\hline
	Israel                                                         & 2                                                                      & 2                                                                      & 1                                                                      & 2                                                                      & 2                                                                      & 2                                                                      & 2                         & 2                           & 1                           & 5                           \\ 
	\hline
	Mongolia                                                       & 1                                                                      & 3                                                                      & 1                                                                      & 1                                                                      & 4                                                                      & 1                                                                      & 1                         & 4                           & 1                           & 5                           \\ 
	\hline
	Fiji                                                           & 0                                                                      & 1                                                                      & 0                                                                      & 2                                                                      & 5                                                                      & 1                                                                      & 1                         & 4                           & 0                           & 5                           \\ 
	\hline
	Uruguay                                                        & 0                                                                      & 2                                                                      & 0                                                                      & 1                                                                      & 3                                                                      & 1                                                                      & 0                         & 2                           & 0                           & 5                           \\ 
	\hline
	Chile                                                          & 0                                                                      & 2                                                                      & 0                                                                      & 1                                                                      & 3                                                                      & 1                                                                      & 0                         & 2                           & 0                           & 5                           \\ 
	\hline
	Unknown                                                        & 1                                                                      & 1                                                                      & 0                                                                      & 3                                                                      & 1                                                                      & 0                                                                      & 1                         & 1                           & 0                           & 4                           \\ 
	\hline
	Zambia                                                         & 1                                                                      & 3                                                                      & 0                                                                      & 1                                                                      & 3                                                                      & 0                                                                      & 1                         & 3                           & 0                           & 4                           \\ 
	\hline
	Azerbaijan                                                     & 1                                                                      & 0                                                                      & 2                                                                      & 1                                                                      & 1                                                                      & 2                                                                      & 1                         & 1                           & 2                           & 4                           \\ 
	\hline
	Guinea                                                         & 0                                                                      & 4                                                                      & 0                                                                      & 0                                                                      & 4                                                                      & 0                                                                      & 0                         & 4                           & 0                           & 4                           \\ 
	\hline
	Georgia                                                        & 1                                                                      & 1                                                                      & 0                                                                      & 2                                                                      & 2                                                                      & 0                                                                      & 2                         & 2                           & 0                           & 3                           \\ 
	\hline
	Mali                                                           & 0                                                                      & 3                                                                      & 0                                                                      & 0                                                                      & 3                                                                      & 0                                                                      & 0                         & 3                           & 0                           & 3                           \\ 
	\hline
	\begin{tabular}[c]{@{}l@{}}Chinese\\Taipei\end{tabular}        & 2                                                                      & 1                                                                      & 0                                                                      & 2                                                                      & 1                                                                      & 0                                                                      & 2                         & 1                           & 0                           & 3                           \\ 
	\hline
	Venezuela                                                      & 0                                                                      & 2                                                                      & 1                                                                      & 1                                                                      & 2                                                                      & 2                                                                      & 0                         & 2                           & 2                           & 3                           \\ 
	\hline
	Paraguay                                                       & 0                                                                      & 3                                                                      & 0                                                                      & 0                                                                      & 4                                                                      & 0                                                                      & 0                         & 4                           & 0                           & 3                           \\ 
	\hline
	Indonesia                                                      & 0                                                                      & 1                                                                      & 1                                                                      & 1                                                                      & 1                                                                      & 2                                                                      & 1                         & 1                           & 2                           & 3                           \\ 
	\hline
	Peru                                                           & 0                                                                      & 1                                                                      & 0                                                                      & 1                                                                      & 2                                                                      & 0                                                                      & 0                         & 2                           & 0                           & 3                           \\ 
	\hline
	Iraq                                                           & 0                                                                      & 3                                                                      & 0                                                                      & 0                                                                      & 3                                                                      & 0                                                                      & 0                         & 3                           & 0                           & 3                           \\ 
	\hline
	Bulgaria                                                       & 1                                                                      & 1                                                                      & 0                                                                      & 1                                                                      & 1                                                                      & 1                                                                      & 1                         & 1                           & 0                           & 3                           \\ 
	\hline
	Lithuania                                                      & 2                                                                      & 1                                                                      & 0                                                                      & 2                                                                      & 2                                                                      & 1                                                                      & 2                         & 1                           & 1                           & 3                           \\ 
	\hline
	Tunisia                                                        & 0                                                                      & 0                                                                      & 0                                                                      & 1                                                                      & 1                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 2                           \\ 
	\hline
	Guatemala                                                      & 1                                                                      & 0                                                                      & 1                                                                      & 1                                                                      & 0                                                                      & 1                                                                      & 1                         & 0                           & 1                           & 2                           \\ 
	\hline
	Bahamas                                                        & 0                                                                      & 2                                                                      & 0                                                                      & 1                                                                      & 2                                                                      & 0                                                                      & 0                         & 2                           & 0                           & 2                           \\ 
	\hline
	\begin{tabular}[c]{@{}l@{}}Trinidad and \\Tobago\end{tabular}  & 0                                                                      & 1                                                                      & 0                                                                      & 1                                                                      & 1                                                                      & 0                                                                      & 0                         & 1                           & 0                           & 2                           \\ 
	\hline
	Montenegro                                                     & 0                                                                      & 1                                                                      & 0                                                                      & 0                                                                      & 2                                                                      & 0                                                                      & 0                         & 2                           & 0                           & 2                           \\ 
	\hline
	Hong Kong                                                      & 0                                                                      & 1                                                                      & 1                                                                      & 0                                                                      & 1                                                                      & 1                                                                      & 0                         & 1                           & 1                           & 2                           \\ 
	\hline
	Moldova                                                        & 0                                                                      & 1                                                                      & 0                                                                      & 0                                                                      & 2                                                                      & 0                                                                      & 0                         & 2                           & 0                           & 1                           \\ 
	\hline
	South Sudan                                                    & 0                                                                      & 1                                                                      & 0                                                                      & 0                                                                      & 2                                                                      & 0                                                                      & 0                         & 2                           & 0                           & 1                           \\ 
	\hline
	Malaysia                                                       & 0                                                                      & 1                                                                      & 0                                                                      & 0                                                                      & 1                                                                      & 0                                                                      & 0                         & 1                           & 0                           & 1                           \\ 
	\hline
	Philippines                                                    & 0                                                                      & 0                                                                      & 0                                                                      & 1                                                                      & 0                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 1                           \\ 
	\hline
	Bahrain                                                        & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 2                                                                      & 0                                                                      & 0                         & 1                           & 0                           & 1                           \\ 
	\hline
	Estonia                                                        & 0                                                                      & 0                                                                      & 1                                                                      & 0                                                                      & 0                                                                      & 1                                                                      & 0                         & 0                           & 1                           & 1                           \\ 
	\hline
	Liberia                                                        & 0                                                                      & 1                                                                      & 0                                                                      & 0                                                                      & 2                                                                      & 0                                                                      & 0                         & 1                           & 0                           & 1                           \\ 
	\hline
	\begin{tabular}[c]{@{}l@{}}United Arab \\Emirates\end{tabular} & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 1                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 1                           \\ 
	\hline
	Jordan                                                         & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 1                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 1                           \\ 
	\hline
	Eritrea                                                        & 0                                                                      & 1                                                                      & 0                                                                      & 0                                                                      & 2                                                                      & 0                                                                      & 0                         & 2                           & 0                           & 1                           \\ 
	\hline
	Singapore                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 1                                                                      & 1                                                                      & 0                         & 0                           & 1                           & 1                           \\ 
	\hline
	Luxembourg                                                     & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 1                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 1                           \\ 
	\hline
	Cyprus                                                         & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 1                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 1                           \\ 
	\hline
	Samoa                                                          & 0                                                                      & 1                                                                      & 0                                                                      & 0                                                                      & 1                                                                      & 0                                                                      & 0                         & 1                           & 0                           & 1                           \\ 
	\hline
	Qatar                                                          & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 0                           \\ 
	\hline
	Tanzania                                                       & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 0                           \\ 
	\hline
	Djibouti                                                       & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 0                           \\ 
	\hline
	Ghana                                                          & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 0                           \\ 
	\hline
	Ivory Coast                                                    & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 0                           \\ 
	\hline
	Tajikistan                                                     & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 0                           \\ 
	\hline
	Kyrgyzstan                                                     & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 1                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 0                           \\ 
	\hline
	Armenia                                                        & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 0                           \\ 
	\hline
	Slovakia                                                       & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 0                           \\ 
	\hline
	Botswana                                                       & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 2                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 0                           \\ 
	\hline
	Grenada                                                        & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 0                           \\ 
	\hline
	Zimbabwe                                                       & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 0                           \\ 
	\hline
	Burundi                                                        & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 0                           \\ 
	\hline
	Kosovo                                                         & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                                                                      & 0                         & 0                           & 0                           & 0                           \\
	\hline
\end{longtable}

\section{Appendix B: Program Codes}
Here are the program codes we used in our research.

% 代码环境示例三则
% 如您的论文不需要展示代码，请删除
% 更多用法，请参考 listings 宏包文档

% Python 代码示例
\begin{lstlisting}[language=Python, name={test.py}]
# Python code example
for i in range(10):
    print('Hello, world!')
\end{lstlisting}

% MATLAB 代码示例
\begin{lstlisting}[language=MATLAB, name={test.m}]
% MATLAB code example
for i = 1:10
    disp("hello, world!");
end
\end{lstlisting}

% C++ 代码示例
\begin{lstlisting}[language=C++, name={test.cpp}]
// C++ code example
#include <iostream>
using namespace std;

int main() {
    for (int i = 0; i < 10; i++)
        cout << "hello, world" << endl;
    return 0;
}
\end{lstlisting}

\end{subappendices}  % 附录内容结束

\end{document}  % 结束